The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-10-25 theoretical-consensus/article/article.tex
add page 05

Due to the self-contained nature of this type of protocols, where the
attestation of the stake is done inside the network, multiple classes
of Proof-of-Stake can be distinguished, by the way the meaningfulness of
the stake is defined and employed. For example, there are the simplest forms of \emph{PoS} that inherit most
of the characteristics of the Proof-of-Work mechanism, but with the
stake as part of the block generation and hashing proof (See Algorithm \ref{alg:PoSBlockGeneration}).
These are commonly referred to as Chain-based \emph{PoS}, inheriting also the
fault-tolerance properties of the Nakamoto consensus protocol.
Committee-based \emph{PoS} protocols, on the other hand, are more complex,
and based on the idea of a committee of entities that are 
selected to lead the block generation and proposal. The committee is
formed by a subset of the network entities, and the selection is done
based on their stake. Usually, the selection takes the form of a random
sequence of stake holders, where the probability of being selected
for every spot is proportional to the stake. That sequence is then 
consciously timed to orderly pace the block generation and proposal, 
by the current slot leader. Examples of this type of mechanisms are
the \emph{Ouroboros} \cite{kiayias2017ouroboros}, \emph{Ouroboros Praos} \cite{david2018ouroboros} protocols
or \emph{Snow White} \cite{daian2019snow}. The longest chain rule is
still applied, with probabilistic finality, and the 1/2 threshold is
still required or tolerated for the network to reach consensus. Nonetheless,
Committee-based \emph{PoS} protocols may not pair well with decentralization and 
increase in the committee size, which detereorates the performance of the
protocol. Several improvements have been proposed to tackle not only this
issue, but also the security of the committees which, smaller, could be more
vulnerable to targeted attacks \cite{david2018ouroboros}.

@@

These aim at achieving deterministic finality, by employing a BFT
consensus protocol, not at the block proposal level, but at the
block acceptance level. The idea is to have a free flow of blocks
proposed by the network, but to have a quorum of entities that
are responsible for the block acceptance. The block proposal mechanism
can be of any kind, for example, Chain-based \emph{PoS}, or Committee-based \emph{PoS},
or even pure Proof-of-Work. The block acceptance layer then allows
for a deterministic acceptance of the blocks, by having a sort of permissioned
BFT consensus protocol, with the trivial 1/3 byzantine tolerance. The most well known examples
and variations are the Tendermint \cite{buchman2016tendermint} and the new Ethereum Casper 
\cite{buterin2017casper} protocols.

Delegated Proof-of-Stake (DPoS) is a variation of the BFT-based protocols,
where the consensus group is formed by a subset of the network entities,
which are selected by the network, with a voting mechanism based on the
delegation of stake. The consensus is then achieved using a typical 
PBFT protocol, with the usual 1/3 byzantine tolerance.

@@

\subsubsection{Proof-of-Elapsed-Time} is another alternative to Proof-of-Work,
especially targetting the intensive mining issue, that has been proposed 
by Intel. The idea is to simulate the mining process by using a
randomized time delay, which is proportional to the amount of work
that would be required to mine the block. The time delay is then used to pace the
block generation and proposal, and the block acceptance can be performed
using a BFT protocol \cite{olson2018sawtooth}. This solution, however,
is not a viable alternative to other, permissioned and permissionless, approaches, 
by holding strong and centralized hardware requirements, 
as the computations require a trusted execution environment
(TEE) to establish the trust among the participants. Some efforts have been made
to combine the TEE feature with Proof-of-Stake, in order to achieve a
more decentralized and secure protocol, but the 
same hardware requirements are still a major issue.

@@

\subsubsection{The XRP Ledger Consensus Protocol} powers the XRP Ledger in a
low-latency, high throughput, and  BFT fashion. The network itself operates somewhat
diferently from the other blockchain references, and treats transactions as atomic units of 
the ledger. These transactions are collected by the nodes and relayed to the
nodes' UNL (Unique Node List), which is a list of peers that are trusted by that current
entity. At every epoch, multiple rounds are needed to filter transactions and vote 
on the next state of the ledger. This voting happens at the 
UNLs' level, which are expected to be overlapped by a certain percentage,
in order to achieve the artificial vote threshold for globally accepting the transactions.
That threshold defines the fault tolerance of the protocol, and is set to 80\%, which 
bounds this protocol to a 1/5 BFT allowance \cite{schwartz2014ripple, chase2018analysis}. 
Despite that limitation, when compared to other BFT protocols, this protocol seeks to trade 
fault tolerance for performance, by having lower connectivity and less message exchanges. 
However, it is still expected high synchrony within the UNLs. Given these characteristics, the protocol
is expected to be more suitable for permissioned networks, and, in terms of network size,
is seen as not as scalable and decentralized as the permissionless ones \cite{survey-dist-consensus}.

>>>>

Due to the self-contained nature of this type of protocols, where the
attestation of the stake is done inside the network, multiple classes
of Proof-of-Stake can be distinguished, by the way the meaningfulness of
the stake is defined and employed. For example, there are the simplest forms of \emph{PoS} that inherit most
of the characteristics of the Proof-of-Work mechanism, but with the
stake as part of the block generation and hashing proof (See Algorithm \ref{alg:PoSBlockGeneration}).
These are commonly referred to as Chain-based \emph{PoS}, inheriting also the
fault-tolerance properties of the Nakamoto consensus protocol.
Committee-based \emph{PoS} protocols, on the other hand, are more complex,
and based on the idea of a committee of entities that are 
selected to lead the block generation and proposal. The committee is
formed by a subset of the network entities, and the selection is done
based on their stake. Usually, the selection takes the form of a random
sequence of stakeholders, where the probability of being selected
for every spot is proportional to the stake. That sequence is then 
consciously timed to orderly pace the block generation and proposal, 
by the current slot leader. Examples of this type of mechanisms are
the \emph{Ouroboros} \cite{kiayias2017ouroboros}, \emph{Ouroboros Praos} \cite{david2018ouroboros} protocols
or \emph{Snow White} \cite{daian2019snow}. The longest chain rule is
still applied, with probabilistic finality, and the 1/2 threshold is
still required or tolerated for the network to reach consensus. Nonetheless,
Committee-based \emph{PoS} protocols may not pair well with decentralization and 
increase in the committee size, which deteriorates the performance of the
protocol. Several improvements have been proposed to tackle not only this
issue, but also the security of the committees which, smaller, could be more
vulnerable to targeted attacks \cite{david2018ouroboros}.

@@

These aim at achieving deterministic finality, by employing a BFT
consensus protocol, not at the block proposal level, but at the
block acceptance level. The idea is to have a free flow of blocks
proposed by the network, but to have a quorum of entities that
are responsible for the block acceptance. The block proposal mechanism
can be of any kind, for example, Chain-based \emph{PoS}, or Committee-based \emph{PoS},
or even pure Proof-of-Work. The block acceptance layer then allows
for a deterministic acceptance of the blocks, by having a sort of permissioned
BFT consensus protocol, with the trivial 1/3 Byzantine tolerance. The most well known examples
and variations are the Tendermint \cite{buchman2016tendermint} and the new Ethereum Casper 
\cite{buterin2017casper} protocols.

Delegated Proof-of-Stake (DPoS) is a variation of the BFT-based protocols,
where the consensus group is formed by a subset of the network entities,
which are selected by the network, with a voting mechanism based on the
delegation of stake. The consensus is then achieved using a typical 
PBFT protocol, with the usual 1/3 Byzantine tolerance.

@@

\subsubsection{Proof-of-Elapsed-Time} is another alternative to Proof-of-Work,
especially targeting the intensive mining issue, that has been proposed 
by Intel. The idea is to simulate the mining process by using a
randomized time delay, which is proportional to the amount of work
that would be required to mine the block. The time delay is then used to pace the
block generation and proposal, and the block acceptance can be performed
using a BFT protocol \cite{olson2018sawtooth}. This solution, however,
is not a viable alternative to other, permissioned and permissionless, approaches, 
by holding strong and centralized hardware requirements, 
as the computations require a trusted execution environment
(TEE) to establish the trust among the participants. Some efforts have been made
to combine the TEE feature with Proof-of-Stake, in order to achieve a
more decentralized and secure protocol, but the 
same hardware requirements are still a major issue.

@@

\subsubsection{The XRP Ledger Consensus Protocol} powers the XRP Ledger in a
low-latency, high throughput, and  BFT fashion. The network itself operates somewhat
differently from the other blockchain references, and treats transactions as atomic units of 
the ledger. These transactions are collected by the nodes and relayed to the
nodes' UNL (Unique Node List), which is a list of peers that are trusted by that current
entity. At every epoch, multiple rounds are needed to filter transactions and vote 
on the next state of the ledger. This voting happens at the 
UNLs' level, which are expected to be overlapped by a certain percentage,
in order to achieve the artificial vote threshold for globally accepting the transactions.
That threshold defines the fault tolerance of the protocol, and is set to 80\%, which 
bounds this protocol to a 1/5 BFT allowance \cite{schwartz2014ripple, chase2018analysis}. 
Despite that limitation, when compared to other BFT protocols, this protocol seeks to trade 
fault tolerance for performance, by having lower connectivity and less message exchanges. 
However, it is still expected high synchrony within the UNLs. Given these characteristics, the protocol
is expected to be more suitable for permissioned networks, and, in terms of network size,
is seen as not as scalable and decentralized as the permissionless ones \cite{survey-dist-consensus}.

\subsubsection{DAG structures} may diverge from the previous approaches with an envisioned 
non-linear structure for the result of the consensus attestations. The choice resides in 
directed acyclic graphs that should expand and physically connect, further in time, the new
data appended to the network. The data, which is typically held in the vertices of the graph, 
can have more or less granularity, either by storing blocks of transactions, or raw transactions directly.
This design decision may pose new potential challenges, related to the verification and finalization of
the consensus protocol. Currently, those two approaches are then distinguished as \emph{blockDAG} and 
\emph{txDAG}. The block-based DAG approach diverges from the traditional blockchain structure by flexibly 
connecting every new block to multiple previous blocks. The rest of the protocol may work in the same way 
as the traditional Proof-of-Work, however, this new feature yields new security considerations. These protocols
are theoretically designed for very high throughput in permissionless environments, but the security trade-offs
are still not well-known.
In a transaction-based DAG, every vertex holds one transaction, that is connected to previous vertices. 
All the sibling vertices shall then represent disjoint transactions to successfully deal with history and conflict
resolutions.
